---
title: "MCRLdata Athena L1 prep"
author: "PR"
date: "2025-05-29"
output: html_document
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)
```

## Purpose

Data downloaded directly from the MCRLdata pipeline via the Athena AWS interface will be cleaned to Level 1 (L1) based solely on flagging. The only decisions being made are if there are multiple sources for a given parameter (e.g. water temperature), picking which dataset is used.

L1 is defined as removing flagged data. Flags are automatically produced by the MCRLdata pipeline, and capture a variety of off-normal data behaviors. I am deciding that L1 will have these values removed, data will be harmonized to a consistent temporal increment, converted to PST, and then exported as a single flat file


### Order of ops

  1. Pull in all datasets
  2. If multiple sources are available, compare and decide which to use
  3. Plot L0 (all data) and L1 to understand what's being scrubbed
  4. Clean to L1, combine, and export
  

## First look

Let's look at what variables are in the files we've downloaded from Athena

```{r set up environment}

source("../scripts/0_setup.R")

```

```{r}
# 2. Discover files ------------------------------------------------------------

## We have several files downloaded as csvs directly from Athena. In an effort to
## make things as plug-and-play as possible, I'm going to add extra steps to ID
## which files are which and read them in, hopefully programmatically
all_files <- list.files("../data/inputs/mcrl_data/athena/210501_250501", 
                        full.names = T)

what_files_do_i_have <- function(file){
  headers <- read_csv(file) %>% 
    # we aren't interested in time or qc, just data cols
    dplyr::select(-c(contains("time"), contains("qc_"))) %>%  
    # make things easier to read, w/ no special characters to potential cause issues
    clean_names() 
  
  paste(colnames(headers), collapse = ", ")
}

file_vars <- all_files %>% 
  map(what_files_do_i_have)

file_vars
```

```{r read in files}

# 3. Read in files -------------------------------------------------------------

## I suspect there's a cleaner way to do this, but it's likely most efficient to
## label files when you're downloading.

## Helper function to find the string you're looking for
which_file <- function(var_string){
  which(sapply(file_vars, grepl, pattern = var_string))[[1]]
}

## One more helper function to clean up column names, round datetime, and convert time to PST
read_csv_formatted <- function(file)(
  read_csv(file) %>% 
    clean_names() %>% 
    mutate(time_utc = round_date(force_tz(time_utc, tzone = "UTC"), unit = "5 min")) %>% 
    mutate(time_pst = with_tz(time_utc, tzone = "Etc/GMT+8"))# %>%  # Forcing to PST
    # mutate(year = year(time_pst), 
    #        quarter = quarter(time_pst, fiscal_start = 3),
    #        month = month(time_pst), 
    #        doy = yday(time_pst), 
    #        hour_of_day = hour(time_pst))
)


## Let's read in our datasets
cdom <- read_csv_formatted(all_files[[which_file("cdom")]]) 
co2 <- read_csv_formatted(all_files[[which_file("p_co2")]]) 
ctd <- read_csv_formatted(all_files[[which_file("salinity")]])
ph <- read_csv_formatted(all_files[[which_file("temperature")]]) ## this is bad regex, need to isolate ph
met <- read_csv_formatted(all_files[[which_file("airtemp")]])
tidegauge <- read_csv_formatted(all_files[[which_file("water_level")]])

```



## Decide on data sources

### Water temp

Water temperature is available from the CTD and pH sensors. While there are clear issues with CTD temperatures for mid 2024 on, the pH temperature does not cover a large-enough range to be useful. 

```{r}
temp_comparison <- full_join(ph %>% 
            filter(qc_temperature == 0) %>% 
            rename("ph_temp" = temperature) %>% 
            dplyr::select(time_pst, ph_temp), 
          ctd %>% 
            filter(qc_temp == 0) %>% 
            rename("ctd_temp" = temp_deg_c) %>% 
          dplyr::select(time_pst, ctd_temp), 
          by = "time_pst")

ggplot(temp_comparison, aes(ctd_temp, ph_temp)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth(se = F)

plot_grid(ggplot(temp_comparison, aes(x = time_pst)) + 
  geom_line(aes(y = ctd_temp), color = "blue", alpha = 0.5), 
  ggplot(temp_comparison, aes(x = time_pst)) + 
  geom_line(aes(y = ph_temp), color = "red", alpha = 0.5), 
  ncol = 1)

# ggplot(temp_comparison, aes(x = time_pst)) + 
#   geom_line(aes(y = ctd_temp), color = "blue", alpha = 0.3) + 
#   geom_line(aes(y = ph_temp), color = "red", alpha = 0.5)
```





